The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-02 manuscript.Rmd
OpenFaaS R template paragraph added

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about containerisation and Rocker Project, and reviewed all sections.
RC \& EH contributed to the section on interfaces for Docker in R (_dynverse_ and `dynwrap`).
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder ready.
HN \& JN co-wrote the section on the T-Mobile use case.
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX contributed content on \pkg{liftr}.
LK contributed content on \pkg{dockr}.
PS wrote the paragraph about OpenFaaS R templates and GitLab CI/CD pipelines with Shinyproxy.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.

>>>>

Parallel efforts to enable R for different cloud providers leads to some level of code fragmentation. This fact and to avoid vendor lock in motivated the [OpenFaaS project](https://www.openfaas.com/), that facilitates the deployment of functions and microservices to Kubernetes or Docker Swarm. OpenFaaS is language agnostic, it provides auto-scaling, metrics, and an API gateway.
Reduced boilerplate code is achieved via templates. R templates\footnote{OpenFaaS R templates: \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}}
are provided based on Rocker's and R-hub's [r-minimal](https://github.com/r-hub/r-minimal) Docker images.
The templates use multi-stage Docker builds\footnote{\herf{https://docs.docker.com/develop/develop-images/multistage-build/}{https://docs.docker.com/develop/develop-images/multistage-build/}} to combine R base images with the OpenFaaS 'watchdog', a tiny Golang webserver that marshals an HTTP request accepted on the API Gateway and invokes the application. 
The R session use \CRANpkg{plumber} or similar packages for the API endpoint with packages and data preloaded, thus minimizing response times.

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about containerisation and Rocker Project, and reviewed all sections.
RC \& EH contributed to the section on interfaces for Docker in R (_dynverse_ and `dynwrap`).
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder ready.
HN \& JN co-wrote the section on the T-Mobile use case.
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX contributed content on \pkg{liftr}.
LK contributed content on \pkg{dockr}.
PS wrote the paragraphs about GitLab CI/CD pipelines with Shinyproxy and OpenFaaS R templates.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.
